Evaluation of probabilistic photometric redshift estimation approaches for \textsc{LSST}

### 2019-03-15 paper/tex/results.tex
reordering section

We begin with a demonstrative visual inspection of the \pzpdf s produced by each code for individual galaxies.
Fig.~\ref{fig:pz_examples} shows the \pzpdf s for four galaxies chosen as examples of \pzpdf\ archetypes: a narrow unimodal redshift solution, a broad unimodal solution, a bimodal distribution, and a multimodal distribution.

@@

\subsection{\Pzpdf\ ensembles}
\label{sec:pitqq}

Fig.~\ref{fig:pitqq} shows a histogram of PIT values, QQ plot, and QQ difference plot relative to the ideal diagonal, showcasing the biases and trends in the average accuracy of the \pzpdf s for each code.
The high QQ values (more high than low PIT values) of \bpz, \cmnn, \delight, \eazy, and \gpz\ indicate \pzpdf s biased low, and the low QQ values (more low than high PIT values) of \skynet\ and \tpz\ indicate \pzpdf s biased high.

@@

Fig.~\ref{fig:pit_stats} displays the values of the KS, CvM, and AD test statistics between the PIT distribution and a uniform distribution $U(0, 1)$, highlighting the relative rather than absolute numbers.
%\red{Can p-values be supplied for each statistic? The statistics themselves are difficult to interpret, other than ``lower is better'' (p-value in skgof was broken, having trouble finding 1-sample KS calculation for uniform distribution)}
\metaphor\ and \lephare\ perform well under the AD but poorly under the KS and CvM due to their high catastrophic outlier rates.
\annz\ and \flexzboost\  are the top scorers under these metrics of the PIT distribution.
\annz's strong performance can be attributed to an aspect of the training process in which training set galaxies with a PIT that more closely matches the percentiles of the DC1 training set's redshift distribution are upweighted; in effect, these quantile-based metrics were part of the algorithm itself that may or may not serve it well under more realistic experimental conditions.

\subsection{Metrics of the stacked estimator of the redshift distribution}
\label{sec:stackedmetrics}

Fig.~\ref{fig:nz} shows the stacked estimator $\hat{N}(z)$ of the redshift distribution for each code compared to the true redshift distribution $\tilde{N}(z)$, where the stacked estimator has been smoothed for each code in the plot using a kernel density estimate (KDE) with a bandwidth chosen by Scott's Rule \citep{Scott:1992} in order to minimize visual differences in small-scale features; the quantiative statistics, however, are calculated using the empirical CDF which is not smoothed.

@@

@@

As expected, \trainz\ perfectly recovers the true redshift distribution: as the training sample is selected from the same underlying distribution as the test set, the redshift distributions are identical, up to Poisson fluctuations due to the finite number of sample galaxies.
\cmnn\ is also in excellent agreement for similar reasons: with a representative training sample of galaxies spanning the colour-space, the sum of the colour-matched neighbour redshifts should return the true redshift distribution.
\flexzboost\ and \tpz\ also perform superb recovery of the true redshift distribution, with only a slight deviation at $z \sim 1.4$.
Our metrics, however, cannot discern whether these four approaches, as well as \delight, are spared the $z \sim 1.4$ degeneracy in $\hat{N}(z)$ because they have more effectively used information in the data or if the impact is simply washed out by the effective average over the test set galaxy sample.
See Appendix~\ref{sec:pointmetrics} for further discussion of the $z \sim 1.4$ issue.

\begin{figure*}
\centering
\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
\caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
We make the observation that these related statistics do not disagree significantly with one another.
\cmnn\ performs comparably well to \trainz, the control case, while \textsc{SkyNet} scores poorly due to an overall bias in its redshift predictions.}
\label{fig:nz_stats}
\end{figure*}

Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
%\red{Can p-values be supplied for each statistic? The statistics themselves are difficult to interpret, other than ``lower is better'' (no, p-values are very difficult to compute for non-uniform distributions)}
\textsc{FlexZBoost}, \textsc{CMNN}, and \textsc{TPZ} outperform the other codes in the $\hat{N}(z)$ metrics.
It is unsurprising that \textsc{CMNN} scores well, as with a near perfectly representative training set means that choosing neighbouring points in color/magnitude space should lead to excellent agreement in the final $\hat{N}(z)$ estimate.
\textsc{TPZ} performed quite poorly in $p(z)$ statistics, but results in a good fit to the overall $N(z)$.
This is somewhat surprising, as performance was optimized for accurate $p(z)$, not $\hat{N}(z)$.
During the validation stage for \textsc{TPZ}, there was a trade off between the width of the $p(z)$ when adjusting a smoothing parameter and overall redshift bias.
The optimal result in the PIT metrics, as illustrated in the shape of the QQ plot, does contain some level of bias as well as a slight underprediction of mean $p(z)$ width, which translates to poor metric scores.
This is something that will be looked into for \textsc{TPZ} in the future.

%\red{discussion of why NN and FZBoost are doing well, how we can try to improve the other codes}

Table~\ref{tab:cdeloss} shows the CDE loss statistic for each photo-$z$ code.
Once again \textsc{FlexZBoost} and \textsc{CMNN} score very well for the stacked $\hat{N}(z)$ metrics, as do \textsc{GPz} and \textsc{TPZ}.
The CDE loss measures how well individual PDFs are estimated, and codes with a low CDE loss tend to have good $\hat{N}(z)$ estimates (though the reverse is not necessarily true).
\textsc{FlexZBoost} is optimized to minimize CDE loss which may explain why the method has good ensemble metrics as well.
Note from Table~\ref{tab:cdeloss} that both \textsc{FlexZBoost} and \textsc{CMNN} have low CDE losses.  %Empirically, we have found that PIT RMSE is not as closely correlated to CDE loss as it is to the $N(z)$ statistics.
As CDE loss is a better measure of individual redshift performance, rather than ensemble distribution performance, this statistic is a better indicator of which codes will be most likely to perform well for science cases where single objects are employed.

%Table~\ref{tab:rmse} gives the root-mean-square-error (RMSE) statistics for both the PIT and N(z) estimators.  The PIT value calculates the RMSE between the quantiles shown in the QQ plot in Figure~\ref{fig:pitqq} and the diagonal, while the N(z) calculates the RMSE between the cumulative distribution of the stacked $\hat{N}(z)$ and the true redshift distribution $N'(z)$.  %\red{more about RMSE?  Or remove it? It's barely described in Section~\ref{sec:metrics} and wasn't even referred to in the text until I added this on June 29th--SJS.}

>>>>

We begin with a demonstrative visual inspection of the \pzpdf s produced by each code for individual galaxies.
Figure~\ref{fig:pz_examples} shows the \pzpdf s for four galaxies chosen as examples of \pzpdf\ archetypes: a narrow unimodal redshift solution, a broad unimodal solution, a bimodal distribution, and a multimodal distribution.

@@

\subsection{Performance on \pzpdf\ ensembles}
\label{sec:pitqq}

Figure~\ref{fig:pitqq} shows a histogram of PIT values, QQ plot, and QQ difference plot relative to the ideal diagonal, showcasing the biases and trends in the average accuracy of the \pzpdf s for each code.
The high QQ values (more high than low PIT values) of \bpz, \cmnn, \delight, \eazy, and \gpz\ indicate \pzpdf s biased low, and the low QQ values (more low than high PIT values) of \skynet\ and \tpz\ indicate \pzpdf s biased high.

@@

Figure~\ref{fig:pit_stats} displays the values of the KS, CvM, and AD test statistics between the PIT distribution and a uniform distribution $U(0, 1)$, highlighting the relative rather than absolute numbers.
%\red{Can p-values be supplied for each statistic? The statistics themselves are difficult to interpret, other than ``lower is better'' (p-value in skgof was broken, having trouble finding 1-sample KS calculation for uniform distribution)}
\metaphor\ and \lephare\ perform well under the AD but poorly under the KS and CvM due to their high catastrophic outlier rates.
\annz\ and \flexzboost\  are the top scorers under these metrics of the PIT distribution.
\annz's strong performance can be attributed to an aspect of the training process in which training set galaxies with a PIT that more closely matches the percentiles of the DC1 training set's redshift distribution are upweighted; in effect, these quantile-based metrics were part of the algorithm itself that may or may not serve it well under more realistic experimental conditions.

\subsection{Performance on the stacked estimator of the redshift distribution}
\label{sec:stackedmetrics}

Figure~\ref{fig:nz} shows the stacked estimator $\hat{N}(z)$ of the redshift distribution for each code compared to the true redshift distribution $\tilde{N}(z)$, where the stacked estimator has been smoothed for each code in the plot using a kernel density estimate (KDE) with a bandwidth chosen by Scott's Rule \citep{Scott:1992} in order to minimize visual differences in small-scale features; the quantiative statistics, however, are calculated using the empirical CDF which is not smoothed.

@@

As expected, \trainz\ perfectly recovers the true redshift distribution: as the training sample is selected from the same underlying distribution as the test set, the redshift distributions are identical, up to Poisson fluctuations due to the finite number of sample galaxies.
\cmnn\ is also in excellent agreement for similar reasons: with a representative training sample of galaxies spanning the colour-space, the sum of the colour-matched neighbour redshifts should return the true redshift distribution.
\flexzboost\ and \tpz\ also perform superb recovery of the true redshift distribution, with only a slight deviation at $z \sim 1.4$.
Our metrics, however, cannot discern whether these four approaches, as well as \delight, are spared the $z \sim 1.4$ degeneracy in $\hat{N}(z)$ because they have more effectively used information in the data or if the impact is simply washed out by the stacked estimator's effective average over the test set galaxy sample.
See Appendix~\ref{sec:pointmetrics} for further discussion of the $z \sim 1.4$ issue.

Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.
%\red{Can p-values be supplied for each statistic? The statistics themselves are difficult to interpret, other than ``lower is better'' (no, p-values are very difficult to compute for non-uniform distributions)}
The stacked estimators of the redshift distribution for \cmnn\ and \trainz\ best estimate $\tilde{N}(z)$ under these metrics, and the only codes that do especially poorly are \eazy, \lephare, \metaphor, and \skynet.
It is unsurprising that \cmnn\ scores well, as with a near perfectly representative training set means that choosing neighbouring points in color/magnitude space should lead to excellent agreement in the final $\hat{N}(z)$ estimate.

\begin{figure*}
\centering
\includegraphics[width=0.74\textwidth]{fig/KSvsCvMvsAD_NZ_withnull_jpg.jpg}
\caption{A visualization of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the $\hat{N}(z)$ distributions.
We make the reassuring observation that these related statistics do not disagree significantly with one another.
\cmnn\ performs comparably well to \trainz, the control case, while \skynet\ scores poorly due to an overall bias in its redshift predictions.}
\label{fig:nz_stats}
\end{figure*}

It is, however, surprising that \tpz\ does well on $\hat{N}(z)$ given its poor performance on the ensemble \pzpdf, especially knowing that \tpz\ was optimized for \pzpdf\ ensemble metrics rather than the stacked estimator of the redshift distribution.
A possible explanation is the choice of smoothing parameter chosen during validation, which affects \pzpdf\ widths as well as overall redshift bias and could be modified to improve performance under the \pzpdf\ metrics.

@@

\subsection{Performance on individual \pzpdf s}

Table~\ref{tab:cdeloss} contains the CDE loss statistic for each code.
Once again \textsc{FlexZBoost} and \textsc{CMNN} score very well for the stacked $\hat{N}(z)$ metrics, as do \textsc{GPz} and \textsc{TPZ}.
The CDE loss measures how well individual PDFs are estimated, and codes with a low CDE loss tend to have good $\hat{N}(z)$ estimates (though the reverse is not necessarily true).
\textsc{FlexZBoost} is optimized to minimize CDE loss which may explain why the method has good ensemble metrics as well.
Note from Table~\ref{tab:cdeloss} that both \textsc{FlexZBoost} and \textsc{CMNN} have low CDE losses.  %Empirically, we have found that PIT RMSE is not as closely correlated to CDE loss as it is to the $N(z)$ statistics.
As CDE loss is a better measure of individual redshift performance, rather than ensemble distribution performance, this statistic is a better indicator of which codes will be most likely to perform well for science cases where single objects are employed.
